Reduced growth due to belowground sink limitation is not fully explained by reduced photosynthesis

### 2016-11-03 manuscript/manuscript_pve.Rmd
added running head
apparently i need to say that the containers had a bottom
edits to abstract removed author contributions manuscript is now double spaced

@@

Sink limitation is known to reduce plant growth, but it is not known how plant carbon (C) balance is affected, limiting our ability to predict growth under sink-limited conditions. We manipulated soil volume to impose sink limitation of growth in *Eucalyptus tereticornis* seedlings. Seedlings were grown in the field in containers of different sizes and planted flush to the soil alongside freely-rooted seedlings. Container volume negatively affected aboveground growth throughout the experiment, but after 120 days, biomass distribution in  leaves, stems and roots was not different. Photosynthetic capacity was significantly reduced in containerized seedlings, and was related to both reduced leaf nitrogen content and starch accumulation. However, the reduction in net leaf photosynthesis over the growth period was insufficient to explain the reduction in growth, so that we also observed an apparent reduction in whole-plant C use efficiency (CUE) with container size. Our results show that sink limitation affects plant growth through feedbacks to both photosynthesis and CUE. Mass balance approaches to predicting plant growth under sink-limited conditions need to incorporate both of these feedbacks.

@@

\
Forty-nine *Eucalyptus tereticornis* seedlings, 20 weeks old and approximately 40 cm tall in tube stock, were chosen from a single local Cumberland plain (western Sydney) cohort. Six additional seedlings were harvested before planting to measure initial leaf area and dry mass of leaves, stems and roots. Six container volumes were used ranging from 5 L to 35 L, with a 22.5 cm diameter, and lengths ranging from 15 to 100 cm. Containers were constructed of PVC pipe, which were capped at the bottom end, and were filled with local top soil. Soil in each container was packed to achieve a target soil bulk density that matched local soil conditions of 1.7 g cm^-3^. An Imidacloprid (BAYER CropScience) insecticide tablet was planted 5 cm below the roots of each seedling. Containers were planted flush with the soil surface inside metal sleeves, designed to minimize excess air space between the container and outside soil while also allowing for container removal. This allowed for soil temperatures in containers to reflect conditions of naturally planted (Free) seedlings. Each experimental block (n=7) contained a complete replicate set of six container volumes as well as one Free seedling, with 1 m^2^ spacing. For each Free seedling, used as the control, a 1 m^2^ subplot was excavated to the hard layer and replaced with the same soil used in each container. A border of root exclusion material was buried 0.25 m deep and extended 0.25 m above the ground surface around each subplot to exclude local vegetation, which was further kept out by periodic weeding.

@@

# Author Contributions
CC contributed to the design of the research, data analysis, collection, interpretation and writing the manuscript. BM contributed to the design of the research, interpretation and writing the manuscript. RD contributed to the design of the research, data analysis, interpretation and writing the manuscript.

>>>>

Runnging Head: Sink limitation and reduced seedling growth

@@

Sink limitation is known to reduce plant growth, but it is not known how plant carbon (C) balance is affected, limiting our ability to predict growth under sink-limited conditions. We manipulated soil volume to impose sink limitation of growth in *Eucalyptus tereticornis* seedlings. Seedlings were grown in the field in containers of different sizes and planted flush to the soil alongside freely-rooted (Free) seedlings. Container volume negatively affected aboveground growth throughout the experiment, and light saturated rates of leaf photosynthesis were consistently lower in seedlings in containers (-26%) compared to Free seedlings. Photosynthetic capacity was also significantly reduced in containerized seedlings, and was related to both reduced leaf nitrogen content and starch accumulation.  After 120 days, harvested biomass of Free seedlings was on average 84% higher than seedlings in containers, but biomass distribution in leaves, stems and roots was not different.
 However, the reduction in net leaf photosynthesis over the growth period was insufficient to explain the reduction in growth, so that we also observed an apparent reduction in whole-plant C use efficiency (CUE) with container size. Our results show that sink limitation affects plant growth through feedbacks to both photosynthesis and CUE. Mass balance approaches to predicting plant growth under sink-limited conditions need to incorporate both of these feedbacks.

@@

\
Forty-nine *Eucalyptus tereticornis* seedlings, 20 weeks old and approximately 40 cm tall in tube stock, were chosen from a single local Cumberland plain (western Sydney) cohort. Six additional seedlings were harvested before planting to measure initial leaf area and dry mass of leaves, stems and roots. Six container volumes were used ranging from 5 L to 35 L, with a 22.5 cm diameter, and lengths ranging from 15 to 100 cm. Containers were constructed of PVC pipe capped at the bottom end and were filled with local top soil. Soil in each container was packed to achieve a target soil bulk density that matched local soil conditions of 1.7 g cm^-3^. An Imidacloprid (BAYER CropScience) insecticide tablet was planted 5 cm below the roots of each seedling. Containers were planted flush with the soil surface inside metal sleeves, designed to minimize excess air space between the container and outside soil while also allowing for container removal. This allowed for soil temperatures in containers to reflect conditions of naturally planted (Free) seedlings. Each experimental block (n=7) contained a complete replicate set of six container volumes as well as one Free seedling, with 1 m^2^ spacing. For each Free seedling, used as the control, a 1 m^2^ subplot was excavated to the hard layer and replaced with the same soil used in each container. A border of root exclusion material was buried 0.25 m deep and extended 0.25 m above the ground surface around each subplot to exclude local vegetation, which was further kept out by periodic weeding.

@@
